---
title: "Predicting County-Level Heart Disease Mortality in the US"
author: "Charlotte Abrams, Laura Cosgrove, Alyssa Vanderbeek"
date: "7 April 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
```


## Introduction
### Background (Alyssa)
### Exploratory Data Analysis (Laura)

## Linear Models

```{r, echo=FALSE}
# load the model fits from previous runs
lm_fit <- readRDS("lm_step_imputed.rds")
ridge_fit <- readRDS("ridge.rds")
mars_fit <- readRDS("mars.rds")
gam_fit <- readRDS("gam_fit_imputed.rds")
lasso_fit <- readRDS("lasso_imputed.rds")
pcr_fit <- readRDS("pcr_imputed.rds")
```


### Stepwise Selection (Alyssa)

### Lasso (Charlotte)

### Ridge (Laura)

### PCR (Charlotte)

## Nonlinear Models

### GAM (Alyssa)

### MARS (Laura)

Note that we chose the most parsimonious model.

## Classification Setting?


## Model Comparison 

### Training RMSE

```{r, echo=FALSE}
set.seed(2)
# resample each model to compare 
res <- resamples(list(
  Stepwise = lm_fit,
  Ridge = ridge_fit,
  Lasso = lasso_fit,
  PCR = pcr_fit,
  GAM = gam_fit,
  MARS = mars_fit
  ))

# plot RMSE ad CI for each model to compare predictive ability
ggplot(res, metric = "RMSE") +
  theme_minimal() +
  labs(title = "Resampled Training RMSE")
```

The more flexible model 

### Test RMSE (Alyssa)

**note: transform test data by centering and scaling with TRAINING means and standard deviations. Does predict do this automatically when using `preProcess` in caret?
-- done

can consider imputation: say 10%, one possible solution. use the training set to build. Look at the preProcess function, can use knn. 
-- done. Still dropped two columns because they had 90\% missing data; imputation not reliable

Check/detect near-zero variance predictors: to decrease computational time and complexity.
-- done. A few of the categories for urban influence came back as having near-zero variance. These were examined as dummy variables; 2000 rows divided across 12 levels means that these are likely to be "rare". We determined that we ought to leave them in the model since they are part of the larger variable "urban_influence".


### Interpretations (Laura)

_Coefficient Shrinkage: Lasso and Ridge_

_Investigation of MARS_

The minimum generalized cross-validation error was achieved

_Investigation of GAM_


## Discussion



\pagebreak
## Appendix

### Data prep

```{r, eval=F}
predictors <- read_csv("./data/Training_values.csv") 
response <- read_csv("./data/Training_labels.csv") 


## Manipulation
data <- response %>% 
  full_join(predictors, by = "row_id") %>%
  separate(col = area__rucc, into = c('metro', 'population'), sep = ' - ') %>%
  rename(urban_influence = area__urban_influence,
         economic_typology = econ__economic_typology) %>%
  mutate(pure_population = fct_collapse(as.factor(population), 
                                        "more_than_1mil" = "Counties in metro areas of 1 million population or more",
                                        "250k_to_1mil" = "Counties in metro areas of 250,000 to 1 million population",
                                        "less_than_250k" = "Counties in metro areas of fewer than 250,000 population",
                                        "more_than_20k" = c("Urban population of 20,000 or more, adjacent to a metro area", 
                                                               "Urban population of 20,000 or more, not adjacent to a metro area"),
                                        "2500_to_20k" = c("Urban population of 2,500 to 19,999, adjacent to a metro area", 
                                                             "Urban population of 2,500 to 19,999, not adjacent to a metro area"),
                                        "less_than_2500" = c("Completely rural or less than 2,500 urban population, adjacent to a metro area", 
                                                        "Completely rural or less than 2,500 urban population, not adjacent to a metro area")),
         economic_typology = as.factor(recode(economic_typology,
                                              "Nonspecialized" = "Nonspecialized",
                                              "Manufacturing-dependent" = "Manufacturing",
                                              "Farm-dependent" = "Farming",
                                              "Federal/State government-dependent" = "Government",
                                              "Mining-dependent" = "Mining",
                                              "Recreation" = "Recreation")),
        metro = factor(metro, 
                       levels = c("Metro", "Nonmetro")),
        urban_influence = str_replace_all(urban_influence, " |/|-", "_"), # replace problematic characters
        urban_influence = str_replace_all(urban_influence, ",", ""), # replace problematic characters
        demo__pct_nonwhite = demo__pct_hispanic + demo__pct_asian + demo__pct_american_indian_or_alaskan_native + demo__pct_non_hispanic_african_american,
        urban_influence = fct_rev(urban_influence),
        metro_adjacency = fct_collapse(population, 
                                       metro = c("Counties in metro areas of 1 million population or more", 
                                                 "Counties in metro areas of 250,000 to 1 million population", 
                                                 "Counties in metro areas of fewer than 250,000 population"),
                                       adjacent = c("Urban population of 20,000 or more, adjacent to a metro area", 
                                                    "Urban population of 2,500 to 19,999, adjacent to a metro area", 
                                                    "Completely rural or less than 2,500 urban population, adjacent to a metro area"),
                                       nonadjacent = c("Urban population of 20,000 or more, not adjacent to a metro area", 
                                                       "Urban population of 2,500 to 19,999, not adjacent to a metro area", 
                                                       "Completely rural or less than 2,500 urban population, not adjacent to a metro area"))) %>% 
  dplyr::select(-demo__pct_hispanic, 
                -demo__pct_asian,
                -demo__pct_american_indian_or_alaskan_native, 
                -demo__pct_non_hispanic_african_american,
                -demo__pct_non_hispanic_white,
                -health__homicides_per_100k, # >90% missing
                -health__pct_excessive_drinking, # >90% missing
                -yr,
                -population) 
```


Training and testing data split. Imputation on missing data.

```{r, eval=F}
## training/test data
set.seed(1)
train_ind <- sample(seq_len(nrow(data)), size = 2/3*nrow(data)) # select rows in 2:1 ratio 

train <- data[train_ind, ] # training dataset
test <- data[-train_ind, ] # testing dataset

# Imputation for missing values with caret, based on training data
training_preproc = caret::preProcess(train, 
                                     method = "knnImpute", # automatically centers and scales data
                                     pcaComp = 10,
                                     na.remove = TRUE,
                                     k = 5,
                                     knnSummary = mean,
                                     outcome = NULL,
                                     fudge = .2,
                                     numUnique = 3,
                                     verbose = TRUE)

# Impute training imputation on both training and testing datasets
train_imputed = predict(training_preproc, train)
test_imputed = predict(training_preproc, test)
```


*Linear Models*

Stepwise regression:
```{r, eval=F}
ctrl = trainControl(method = "repeatedcv", number = 10, repeats = 5)
set.seed(2)
step.fit = caret::train(heart_disease_mortality_per_100k ~ ., 
                        data = train_imputed, 
                        method = 'glmStepAIC',
                        metric = "RMSE",
                        trControl = ctrl)
```


Lasso:
```{r, eval=F}
set.seed(2)
lasso_fit <- caret::train(heart_disease_mortality_per_100k ~ ., 
                          data = heart,
                          method = "glmnet",
                          metric = "RMSE",
                          tuneGrid = expand.grid(alpha = 1,
                                                lambda = exp(seq(-10, 0, length = 200))),
                          preProcess = c("zv"),
                          trControl = ctrl)
```


Ridge:
```{r, eval=F}

```


PCR:
```{r, eval=F}
set.seed(2)
pcr_fit <- caret::train(heart_disease_mortality_per_100k ~ ., 
                        data = heart,
                        method = "pcr",
                        trControl = ctrl,
                        metric = "RMSE",
                        tuneLength = 200)
```



*Non-linear models*


GAM:
```{r, eval=F}
set.seed(2)
gam.fit <- caret::train(heart_disease_mortality_per_100k ~ ., 
                        data = train_imputed,
                        method = "gam",
                        metric = 'RMSE',
                        tuneGrid = data.frame(method = "GCV.Cp", select = c(TRUE, FALSE)),
                        trControl = ctrl)
```


MARS:
```{r, eval=F}

```

