---
title: "Project Proposal"
author: "Laura Cosgrove"
date: "3/7/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Project Title
Predicting Heart Disease in U.S. Counties

## Data Sources (Charlotte)

## Description of Data (Alyssa)

```{r echo = FALSE, message = FALSE}
training_values = read_csv("./data/Training_values.csv")
training_labels = read_csv("./data/Training_labels.csv")
test = read_csv("./data/Test_values.csv")

str_remove_hlth <- function(x) {
  str_remove(x, pattern = "health__")
}

test_health_subset = test %>% 
  select(starts_with("health")) %>% 
  rename_all(str_remove_hlth) %>% 
  skimr::skim_tee()
```

There are variables in four general categories:

(a) Area

(b) Health

(c) Demography

(d) Economic

Of course, each category is related to one another, but we can split them up for the purposes of description.


## Planned Analyses (Laura)

_Part 1: Linear Regression and its Relatives_
We will randomly partition the data into a training and test set (2:1). We plan to fit a ordinary least squares model using backward stepwise regression, a lasso model, a ridge model, and a PCR model to these data. We will select the lasso, ridge, and PCR parameters using repeated cross-validation. The model performance will be compared using the test RMSE.

_Part 2: Nonlinear Regression_
We will conduct a similar analysis using nonlinear methods. Using predictor variables suggested by the best-performing model in part 1, we will fit smoothing splines if we detect a potential nonlinear trend in the outcome variable, with roughness parameter chosen by LOOCV. We will evaluate the performance of the spline model vs. the linear model using the test RMSE.